---
title: "eBay Homework"
output:
  html_document:
    df_print: paged
---
1: That women bid higher than men
2: That bidders will pay more for a new item than an old one

To do that we will examine data from eBay auctions (and look at some prior literature). The dataset contains the results from auctions completed in 2008 for Apple 3rd Gen iPods (1GB). 

Instructions:

1. Find, download, and read the following article:
Ba, S., & Pavlou, P. A. (2002). Evidence of the effect of trust-building technology in electronic markets: Price premiums and buyer behavior. MIS Quarterly, 243-268.

Pay special attention to Study 2

2. Create an R Notebook Titled eBay Homework

3a. Read in the following eBay data:

```{r}
library(tidyverse)
exfile = "http://www.itk.ilstu.edu/faculty/jrwolf/ebay.csv"
df = read_csv(exfile)
head(df)
```


# b. Examine the data (look at head and tail)
```{r}
head(df)
```
```{r}
tail(df)
```

# c. Examine the structure (str) of the data
```{r}
str(df)
```

# d1. Find the Mean (mean) and Standard Deviation (sd) by old/new of Price
```{r}
oldPrices = df %>%
  filter(Used == 1) %>%
  summarize(meanUsed = mean(Price, na.rm = TRUE), sdUsed = sd(Price, na.rm = TRUE))
oldPrices
newPrices = df %>%
  filter(New == 1) %>%
  summarize(meanNew = mean(Price, na.rm = TRUE), sdNew = sd(Price, na.rm = TRUE))
newPrices


```

# d2. Find the Mean (mean) and Standard Deviation (sd) by male/female of Price
```{r}
maleBuys = df %>%
  filter(Gender == "M") %>%
  summarize(meanUsed = mean(Price, na.rm = TRUE), sdMale = sd(Price, na.rm = TRUE))
maleBuys
femaleBuys = df %>%
  filter(Gender == "F") %>%
  summarize(meanNew = mean(Price, na.rm = TRUE), sdFemale = sd(Price, na.rm = TRUE))
femaleBuys
```

# e1. Create a boxplot of Price (by old/new )
```{r}
df %>%
  ggplot(aes(x = Cond, y = Price)) +
  labs(title = "Price Variation", y = "Prices", x = "Old and New") +
  geom_boxplot()
```

# e2. Create a boxplot of Price (by male/female )
```{r}
df %>%
  ggplot(aes(x = Gender, y = Price)) +
  labs(title = "Price Variation", y = "Prices", x = "Male and Female") +
  geom_boxplot()
```

# f. Create a histogram of Price (by color)
```{r}
df %>%
  group_by(Color) %>%
  ggplot(aes(Price)) +
  labs(title = "Histogram of Price",
       x = "Price Range",
       y = "Price Value") +
  geom_histogram(binwidth = 4) +
  facet_wrap(~Color, ncol = 3)
```

# g1. Run a t-test on Price (by old/new)
```{r}
t.test(Price ~ Cond, data = df)
```

# g2. Run a t-test on Price (by male/female)
```{r}
t.test(Price ~ Gender, data = df)
```

# h. report your findings for both t-tests using APA format 

There is a big statistical difference, when we check the Condition of the product whether it is new item price (M = 42.79, SD = 8.77) compared to the Old item price(M = 31.40, SD = 11.18), t(790) = 17.9, p < 2.2e-16.

On average, Women spent more on items (M = 38.49) compared to Men(M = 37.83) this difference was not at all significant, t(1044) = 0.95, p = 0.34.
